---
title: "Solow Model - Assumptions, Results and Steady-State Dynamics"
author: "Sam Drew"
date: "18/11/2016"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction

The Solow model is an extremely simple economic model, introduced by Robert Solow in 1956, which explores the relationship between capital and effective labour in a market economy. In doing so it provides results relating to the expected growth rates of a nation, the capital share of labour, labour wage growth and capital growth. 

This essay will build up the model step-by-step, exploring the assumptions implied by each element of the model implications and the limitations of these results as implied by the assumptions.

# The Model

The model states that output ($Y$) can be described as a function of capital ($K$) and effective labour, described by two parameters, the labour force ($L$) and the labour efficiency, or technology factor ($A$).

$$Y = F(K,AL)$$

where capital growth is defined as 

$$\dot K = s Y - \delta K$$

with the exogenous rates for savings ($s$) and depreciation ($\delta$). 

This will model a closed market economy, as it has no trade. Additionally the capital increase is proportional to the savings rate, because the savings and investment are assumed to be equal.

## Analytical model

In order to make use of this model, it is necessary to give it mathematical form, which will further encapsulate the behaviour we want to represent in this model. In this case we are expecting diminishing, but positive returns from both capital and effective labour. An effective way of representing this is as a Cobb-Douglas function.

$$Y = K^\alpha(AL)^\beta$$

Given the additional assumption that in order to double the output you will need to double both the capital stock and effective labour, we can conclude that $\alpha + \beta = 1$ therefore $\beta = 1 - \alpha$. This assumption implies the economy is large enough to already be considered efficient such that additional competition, and that the growth factors for $A$ and $L$ are exogeneous, so neither technological growth nor immigration rates will be effected by output. 

$$\begin{aligned}
L(t) = L(0)e^{nt}\\
A(t) = A(0)E^{gt}
\end{aligned}$$

from which we show that the growth rates ($n$ and $g$) are defined as the 

$$\begin{aligned}
\frac{\dot L}{L} = \frac{dL(t)/dt}{L(t)} = \frac{nL(0)e^{nt}}{L(0)e^{nt}} = n \\
\frac{\dot A}{A} = \frac{dA(t)/dt}{A(t)} = \frac{nA(0)e^{nt}}{A(0)e^{gt}} = g 
\end{aligned}$$.

## Analysis

More interesting than looking at the capital output for the economy is looking at the output per unit of effective labour. This can be achieved just by dividing through by AL, which gives 

$$\frac{Y}{AL} = F(\frac{K}{AL},1)$$

which we can then denote as 

$$y = f(k)$$

and

$$ \dot k = s \cdot f(k) - (n + g + \delta)k $$

where $y=\frac{Y}{AL}$, $k=\frac{K}{AL}$. If we then assume that capital per unit effective labour is constant (i.e. capital growth per worker is zero: $\dot k = 0$) we get

$$\begin{aligned}
0 &= s \cdot f(k) - (n + g + \delta)k \\
\Rightarrow s \cdot f(k) &= (n + g + \delta)k
\end{aligned}$$

Graphically this can be plotted as 

```{r, echo=FALSE}
xsize = c(0,10)
ysize = c(0,3)
n = 0.05
g = 0.1
d = 0.1
curve(x^0.3, axes = FALSE, frame.plot = TRUE, xlab = "k (Capital/EffLabour)", ylab = "y (Output/EffLabour)", xlim = xsize, ylim= ysize)
par(new = TRUE)
curve((n + g + d)*x, axes = FALSE, xlab = "", ylab = "", xlim = xsize, ylim = ysize)
```

which gives us an equilibrium point for labour share of output with respect to capital $k^*$.


