---
title: "MomentumPortfolioTheory"
author: "Sam Drew"
date: "25/01/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
require("quantmod")
require("xts")
require("zoo")
```

# Theory

The theory is that if you take the top quantile of performing stocks purchase them, while at the same time selling the bottom quantile, the outcome is short-term profit. I want to test this, so let's get hold of FTSE 250 data. Earlier I got hold of all the stock symbols for ftse250 companies, using a spider/scraper.

```{r loadData, include=FALSE}
#Load Data
ftse250 <- read.csv("shares.csv",as.is = TRUE)

makeLondon <- function(x) {
  if (endsWith(x, ".")) {
    return(paste(x, "L", sep="") )
  } else {
    return(paste(x, ".L", sep=""))
  }
}

setSymbolLookup(CTY.L='google',LCL.L='google') #Yahoo 404'd on these stocks
ftse250$symbol <- sapply(ftse250$symbol, makeLondon)
Data<-new.env()
quantmod::getSymbols(ftse250$symbol)
sets <- ls(pattern = "\\.L$")
```

That done we now want to merge the data. Or not. It broke my computer.

```{r mergeData, include=FALSE}
setlist <- mget(sets)
openVal <- function(symbol) {
  return(paste(symbol,'.Open',sep = ""))
}

pcchange <- function(ts, values) {
  values <- coredata(values)
  size <- dim(values)[1]
  retval <- array(dim = size)
  for(i in ts+1:size) {
    retval[i] <- (values[i] - values[i-ts])/values[i]
  }
  return(retval[1:size])
}

#Should have used rollapply (zoo)
largest = 0
largestset = ""
for (set in sets) {
  # Get the last value from the dataset and divide through and calculate momentum
  size <- dim(setlist[[set]])[1]
  if (size > largest) {
    largest = size
    largestset = set
  }
  base <- drop(coredata(setlist[[set]][dim(setlist[[set]])[1],paste(set,".Open", sep = "")])) 
  setlist[[set]]$rebase <- setlist[[set]][,1] / base
  setlist[[set]]$pcchange1d <- pcchange(1, setlist[[set]][,1])
  setlist[[set]]$pcchange1d <- na.fill(setlist[[set]]$pcchange1d,0)
  setlist[[set]]$rollmean <- rollmeanr(setlist[[set]]$pcchange1d, 20)
  attributes(setlist[[set]])$sd <- sd(setlist[[set]]$rollmean, na.rm=TRUE)
}
```

So now that I have some data sets, I need to compare them, and set up the behaviour...

```{r compareData, include=FALSE}

mergedRollmean <- setlist[[sets[1]]]$rollmean
mergedOpen <- setlist[[sets[1]]]
for (set in sets[-1]){
  mergedRollmean <- cbind.xts(mergedRollmean, setlist[[set]]$rollmean)
}
colnames(mergedRollmean) <- sets

meanquantiles <- apply.daily(mergedRollmean, function(x) quantile(x, prob=c(0.2, 0.5,0.8), na.rm=TRUE))
#mergedRollmean <- cbind(mergedRollmean, meanquantiles)

#TOOD Optimize this code - far too slow at the moment
behaviour <- apply.daily(mergedRollmean, function(x){
  index = as.Date.POSIXct(attributes(x)$index)
  #print(x)
  noQuantiles <- (is.na(meanquantiles[index]$`80%`) | is.na(meanquantiles[index]$`20%`))
  return(sapply(x, function(x) {
    if (noQuantiles | is.na(x)) return(NA)
    else if (x >= meanquantiles[index]$`80%`) return(TRUE)
    else if (x <= meanquantiles[index]$`20%`) return(FALSE)
    else return(NA)
  }))
})
```

Now that the portfolio behaviour has been defined, we'll take a look at the 3-month returns for hold and short positions, and composite. We'll then compare these to the index results.

```{r calculateReturns, include=FALSE}
time <- 91 #Number of days for 3 months

apply.daily(behaviour, function(x) {
  index = as.Date.POSIXct(attributes(x)$index)
  return(sapply(x, function(y) {
    if (is.na(x)) {
      return(0)
    } else if (x) {
      name = attributes(y)$dimname[[2]]
      priceToday <- coredata(setlist[[name]][index,1])
      price3months <- coredata(setlist[[name]][index+time,1])
      return((price3months - priceToday)/ priceToday)
    } else {
      name = attributes(y)$dimname[[2]]
      priceToday <- coredata(setlist[[name]][index,1])
      price3months <- coredata(setlist[[name]][index+time,1])
      return((priceToday - price3months)/ priceToday)
    }
  }))
})
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
